---
title: "HW2"
author: "Canyon Foot"
date: "2/9/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(babynames)
library(tidyverse)
library(pnwflights14)
library(dplyr)
library(gridExtra)
library(pnwflights14)
Pres <- read_excel("~/canyon-foot/homework/Pres.xlsx")
Pres_list <- read_excel("~/canyon-foot/homework/Pres_list.xlsx")
data("Babynames")
data("flights")
data("weather")
data("planes")

```

1)



```{r}
presidential_names <- c("Grover","Benjamin","Grover","William","Theodore","William","Woodrow","Warren","Calvin","Herbert","Franklin","Harry","Dwight","John","Lyndon","Richard","Gerald","James","Ronald","George","William", "Barack")
baby_pres <- babynames %>% 
  group_by(year) %>% 
  full_join(Pres) 
baby_pres <- baby_pres %>% 
  full_join(Pres_list) %>%
  mutate(ContemporaryPresTF = (name == president | name == prev_pres | name == next_pres), OverallPresTF = name %in%  c(presidential_names)) %>% 
  filter(sex == "M")

baby_pres <-baby_pres %>% 
  filter(year > 1880, ContemporaryPresTF == TRUE & OverallPresTF == TRUE)

#This code combines my data into one data set and filters for names of presidents. 
# It then filters for names that were the name of the current president, the previous
# president or the next president. This is how I got the ranges of years that
# each name is displayed for.

```

```{r, fig.width= 12}
plot_1960 <-baby_pres %>% 
  filter(prop < .0075 & year < 1960) %>% 
  ggplot(aes(x = year, y = prop, color = name)) + geom_line(size = 2) + labs(x = "Year", y = "Name Proportion", title = "Presidential Name Popularity 1882 - 1960                                   Black Bars Indicate End of a Presidency") +
  #making first line plot
  annotate("linerange", x = c(1884, 1888, 1892, 1900, 1912, 1920, 1923, 1928, 1932,  1945, 1952), ymin = 0, ymax = .0008) +
  
  annotate("text", x = c(1886, 1894, 1905, 1905, 1916, 1926, 1935, 1939,1950, 1955), y = c(.0023, .005, .0035, .001, .006, .0015, .0067, .0015,.0035, .0008), label = c("Grover", "Benjamin", "Theodore", "Woodrow", "Warren", "Calvin", "Herbert", "Franklin", "Harry", "Dwight"), size = 6) +
  annotate("text", x = 1956, y = .007, label = "Note: William Taft and \n William Mikinley excluded \n for scale reasons") +  theme(plot.title = element_text(size=22))
# annotating for presidental change and adding names
plot_2015 <- baby_pres %>%
  filter(year >= 1960, name != "Dwight" & name != "Theodore" & name != "Harry" & name != "Benjamin" & (name != "Ronald" | year < 1992)) %>% 
  
ggplot(aes(x = year, y = prop, color = name)) +
  geom_line(size = 2) +
  labs(x = "Year", y = "Name Proportion", title = "Presidential Name Popularity 1960 - 2015") +
   #making second line plot
  annotate("linerange", x = c(1960, 1963, 1968, 1974, 1976, 1980, 1988, 1992,2000,2008), ymin = 0, ymax = .0045) +
  
  annotate("text", x = c(1964, 1968, 1985, 1998), y = c(.035, .015, .022, .007), label = c("John", "Richard", "James", "William"), size = 6)  +
  
  annotate("text", x= c(1968, 1978, 1985, 2000, 2010), y = -.001, label = c("Lyndon", "Gerald", "Ronald", "George", "Barack"), size = 6) + theme(plot.title = element_text(size=22))
# annotating for presidental change and adding names
```

Prior to the 1960s, there was usually a strong spike in the popularity of the president's name around his election. After the 1960s this effect  essentially dissapeared.
```{r, fig.height= 15, fig.width= 15}
grid.arrange(plot_1960, plot_2015, nrow = 2)

# combining plots into one image
```

Book Exercises:
4.1

1) Summarise
2) Mutate
3) Arrange
4) filter
5) Select

4.2
January had the highest proportion of cancelled flights, October had the lowest. This makes intuitive sense as the fall tends to have less severe weather than the winter.
```{r}
flights %>% group_by(month) %>% 
  mutate(cancelled = if_else(is.na(air_time),1,0)) %>%
  summarise(cancelled_rate = sum(cancelled)/n()) %>%
  arrange(desc(cancelled_rate))
```


4.3
The plane with tail number N223AG flew the most flights from the pnw.
```{r}
flights %>% group_by(tailnum) %>% 
  summarise(total = n()) %>% 
  arrange(desc(total))

flights %>% filter(tailnum == "N223AG") %>% 
  group_by(month) %>% summarise(total = n()) %>% 
  ggplot(aes(x= month, y = total)) + geom_line()
```

4.4
The oldest plane has tailnum "N201AA". There is a total of 2628 planes listed in the planes table.
```{r}

planes %>% filter(year > 0) %>% 
  full_join(flights, "tailnum")  %>%
  arrange(year.x)
```

4.5
61 planes have missing years of manufacture. Boeing, Airbus Industrie, Airbus, Bombardier Inc, and McDonnel Douglas are the most popular manufacturers.

For all manufacturers, production was highest around the year 2000. Since then, Boeing has dominated the market, with AirBus a fairly distant second.
```{r}
planes %>% filter(year == 0 | is.na(year))
# planes without year
planes %>% group_by(manufacturer) %>% summarise(total=n_distinct(tailnum)) %>% arrange(desc(total))

planes %>% filter(year > 1960, !is.na(manufacturer)) %>% 
  group_by(manufacturer, year) %>% 
  summarise(total=n_distinct(tailnum)) %>% 
  mutate(manufacturer2 = if_else(!(manufacturer %in% c("AIRBUS", "AIRBUS INDUSTRIE", "BOEING", "BOMBARDIER INC", "MCDONNELL DOUGLAS")), "Other", manufacturer)) %>%
  
  # creating other catagory
  ggplot(aes(x = year, y = total, color = manufacturer2)) +
  geom_line()


```

The temperature distribution is centered around 65 degrees, and is right skewed.

The relationship between dew point and humidity is positive and fairly weak, the relationship between visibility and precipitation is negative and moderate.
```{r}
weather %>% filter(month == 7) %>% 
  ggplot(aes(x = temp)) + geom_histogram(binwidth = 5)
# histogram
weather %>% filter(month == 7) %>% 
  ggplot(aes(x = dewp, y = humid)) + 
  geom_jitter() +
  geom_smooth(se = FALSE, method = "lm")

weather %>% filter(month == 7, precip< .10) %>% 
  ggplot(aes(x = precip, y = visib)) + 
  geom_jitter() + 
  geom_smooth(se = FALSE, method = "lm")
# scatterplots
```





