---
title: "Activity-3"
author: "Canyon Foot"
date: "2/6/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(babynames)
data("babynames")

```
The least popular popular names are shown in the table below.
```{r}
babynames %>%
  mutate(is_popular = prop > .01) %>%
  filter(is_popular == TRUE) %>%
  arrange(prop)
```
Bella was most common in the late 2000s and early 2010s. I wonder if there was a popular series of books and movies featuring a main character named "Bella" that was released during this period...

The class of Bella is "tbl_df"
```{r}
bella <- babynames %>% filter(name == "Bella" & sex == "F") %>% select(year,sex,name,n)
ggplot(bella, aes(x = year, y = n)) + geom_line()

class(bella)

bella %>% arrange(desc(n))
```
My name was most popular in 2003, with n = 158. The most popular name that year was Jacob.
```{r}

babynames %>% filter(name == "Canyon" & sex == "M") %>% arrange(desc(n))
babynames %>% filter(year == 2003) %>% arrange(desc(n))

babynames %>% filter(name == "Canyon" & sex == "M") %>% arrange(desc(n)) %>% ggplot(aes(x = year, y = n)) + geom_line()
```
1957 had the highest number of Births.
```{r}
bab <- babynames %>% group_by(year) %>% summarise(count = sum(n))
bab %>% arrange(desc(count))

babynames %>% group_by(name) %>% summarise(early = min(year), late = max(year))
```
I does appear that recently there was a fairly sharp increase in the number of distinct names. This effect seems somewhat stronger for female names.
```{r}
babynames %>% group_by(year, sex) %>% mutate(num_names = n_distinct(name)) %>% 
  ggplot(aes(x = year, y = num_names, color = sex)) + geom_line()

```
The most popular names of the 1990s were Andrew, Ashley, CHristopher, Daniel, Jacob, Jessica, Joshua, Matthew, Michael, and Nicholas.
```{r}
babynames %>% filter(year >= 1990, year < 2000, prop > .0005) %>% group_by(name) %>% mutate(total_prop = sum(prop)) %>% filter(total_prop > .13) %>%
  ggplot(aes(x = year, y = n, color = name, linetype = sex)) + geom_line()
```

```{r}
Men_of_The_Bible <- c("Aaron", "Abel", "Abraham", "Adam", "Andrew", "Asher","Barak","Barnabas","Bartholomew","Benjamin","Daniel","Elijah","Elon","Ephraim",
"Ezekiel","Gabriel","Gad","Gideon","Hillel","Hiram","Immanuel","Isaac","Ishmael",
"Israel","Jabin","Jacob","James","Jared","Jesse","Jesus","Jethro","Joel","John","Jokim","Joseph","Joshua","Josiah","Judah","Levi","Lucas","Luke","Mark","Matthew","Meshach","Micah","Moses","Noah","Paul","Peter","Philip","Reuben","Rufus","Samson","Saul","Seth","Silas","Simon","Solomon","Stephen","Thaddeus","Thomas","Timothy", "Zacharias")

Women_of_the_Bible <- c("Abigail", "Ada", "Ahlai", "Angel", "Anna", "Apphia", "Atarah", "Athaliah", "Bathsheba", "Bilhah", "Candace", "Chloe", "Claudia", "Damaris", "Deborah", "Delilah", "Diana", "Dinah", "Elizabeth", "Esther", "Eunice", "Eve", "Hagar", "Hannah", "Huldah", "Jedidah", "Jezebel", "Joanna", "Judith", "Julia", "Leah", "Lois", "Lydia", "Martha", "Mary", "Michaiah", "Milcah", "Miriam", "Naarah", "Naomi", "Orpah", "Phebe", "Priscilla", "Rachel", "Rebecca", "Ruth", "Salome", "Sapphira", "Sarai", "Sherah", "Susanna", "Tabitha", "Tamar", "Tirza", "Vashti", "Zilpah")

# To view the popularity of Biblical names I created two vectors, "Women_of_the_Bible" and "Men_of_the_Bible" that each contain ~50 of the most common Biblical names, then filtered for names included in these vectors. 
babynames %>%
  filter((sex == "M" & name %in% Men_of_The_Bible) | (sex == "F" & name %in% Women_of_the_Bible)) %>%
  group_by(year, sex) %>% 
  mutate(total_prop = sum(prop)) %>%
  ggplot(aes(x = year, y = total_prop, color = sex)) + geom_line() + labs(x = "Year", y = "Total Proportion of Biblical Names", title  = "Use Of Biblical Names Over Time")


# Tracking Pppular Biblical names over time. I used "sum(prop) > .007 * 135" to find the names with the highest average proportion over the 135 years of the data. Note that this may exclude popular biblical names that were only popular for a short period of time.

babynames %>% group_by(name) %>%
  filter((sex == "M" & name %in% Men_of_The_Bible) & sum(prop) > .007 * 135 | (sex == "F" & name %in% Women_of_the_Bible & sum(prop) > .007 * 135)) %>%
  group_by(year) %>% 
  ggplot(aes(x = year, y = prop, color = name)) + geom_line() + labs(x = "Year", y = "Proportion", title  = "10 Most Popular Biblical Names Over Time")
```







